Controlling the dewetting morphologies of thin liquid films by switchable substrates

### 2023-04-21 main_shortened.tex
New replies, thanks for nothing ref A

\bibstyle{apsrev}
\usepackage{bm}%
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}

@@

\begin{abstract}
\noindent Switchable and adaptive substrates emerged as valuable tools for the control of wetting and actuation of droplet motion. Here we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modelled with a space and time varying contact angle.
With a static pattern, all the fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed. 
A criterion discriminating whether rivulets can be observed or not is identified in terms of a single dimensionless parameter.
Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
\end{abstract}

@@

\noindent {\it Introduction}. Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,DERYCK1998278,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,jo2009evaluation,Wijshoff2010}. 
Moreover, the continuously growing technological interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics~\cite{Kim_2005, Luechinger_2008} and printable photovoltaics~\cite{Brabec2001,RH20}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018,GuoGuo_RSCAdv2016}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{LiuChem2005, XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010,ayres2007stimuli,YongEtAl_Mat2018}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000,DelormeEtAl_Langmuir2005,OscuratoEtAl_AAMI2017} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
While a consistent body of theoretical/computational work has been devoted to processes on static heterogeneous substrates, though, the time dependent case is still almost unexplored, with few relevant exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to an analysis of the linear regime~\cite{suman2006dynamics}.

In this Letter, we study, by means of numerical simulations, the full dewetting dynamics of a thin liquid film deposited over a substrate with a time varying wettability pattern, from rupture to the long time morphology. 
We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on the static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases. These observations are then explained theoretically.
We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, it is possible to control, to some extent, the dewetting morphology. In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
We introduce a control parameter that discriminates whether rivulets or just droplets (as in the static situation)
can be observed and we propose a phenomenological argument that justifies the logarithmic dependence of the rivulets life time on the pattern speed.\\

@@

\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{ReynoldsLubr,RevModPhys.69.931,RevModPhys.81.1131} 

@@

by means of a recently developed method, based on a lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.

@@

The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$~\footnote{\textcolor{red}{The Laplace pressure is the product of the surface tension 
and the local curvature, therefore its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$; 
since we are in the lubrication regime, though, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 

@@

field~\cite{Benet2014,juanes2018}.}}, 
\textcolor{red}{where $\gamma$ is the surface tension}.

@@

\textcolor{red}{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))

@@

\textcolor{red}{where $f(\xi)=\xi^{-3} - \xi^{-2}$} and $h_{\ast}$ is the height at which the disjoining pressure vanishes and sets the precursor layer thickness~\footnote{\textcolor{red}{Expressions of 
the general form $\Pi(h) = \gamma(1-\cos \theta)(n-1)(m-1)/(n-m)((h_{\ast}/h)^n - (h_{\ast}/h)^m)/h_{\ast}$
are frequently found in the literature, with various combination of exponents $(n,m)$ ($n>m$), depending
on the type of interactions among fluid molecules and between fluid and substrate~\cite{SCHWARTZ1998173,Mitlin,Teletzke}. A common choice, for instance, is the pair $(9,3)$, which 
can be derived from molecular interactions of Lennard-Jones type. Here, the use of the $(3,2)$ pair is motivated mainly by the computational advantage of keeping the characteristic dewetting time 
shorter. We have run also a single case ($\lambda=256 h_0$ and $\Gamma=1.5$) with $(9,3)$ and observed that 
the dynamics remains essentially the same, but for a shift in the time scale (see also~\cite{PhysRevE.104.034801})}}.
The time variation of the patterned substrate then enters the model precisely through the disjoining pressure, by making the contact angle space and time dependent, i.e. $\theta = \theta(\mathbf{x},t)$.
\textcolor{red}{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
directions~\cite{KarguptaLangmuir2003,Jalali2018,Nagayama2020,Das2020}.}

@@

% \textcolor{red}{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 

@@

We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Notice that, since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one LB time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by one $\Delta x$ every $1/v_{\theta x}$ time steps (and equivalently in the $y$-direction).}.
\textcolor{red}{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film \cite{Mecke_2005}. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$ \cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 

@@

    \includegraphics[width=0.45\textwidth]{Figure_1.png}

@@

    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values coded in light blue (yellow).

@@

\textcolor{red}{To give a flavour of how to compare with actual experimental systems, let us notice
first that the expression of $t_0$ depends on the disjoining pressure (through its dependence 
 $q_0 = \sqrt{\frac{1}{2\gamma} \Pi^{\prime}(h_0)}$) and contains, therefore, energetic details, namely 
 the Hamaker constant $A_H$. Consider, for instance, the problem studied in \cite{becker2003complex}: the dewetting of a polystyrene film of thickness $h_0 \approx 4 \, \text{nm}$ deposited on a silicon 
dioxide substrate, where $A_H  = 2.2 \times 10^{-20} \, \text{J}$ and 
$\gamma = 0.03 \, \text{N}/\text{m}$. Neglecting the short range part of the interface potential
(such that $\Pi(h)= - \frac{A_H}{6\pi h^3}$)~\cite{Mecke_2005,becker2003complex}, one gets 
$q_0^2 \approx \frac{A_H}{4\pi \gamma h_0^4} \approx 2.5 \times 10^{-4} \, \text{nm}^{-2}$.
The characteristic time of the growth of the instability for these experiments then reads $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4} \approx 300 \, \text{s}$ (the dynamic viscosity of polystyrene is 
$\mu = 1.2 \times 10^4 \, \text{Pa} \cdot \text{s}$)~\cite{PhysRevLett.99.114503}. It can be seen that indeed film rupture occurs within time 
scales of the order of $t_0$~\cite{becker2003complex}, as in our 
simulation for the largest pattern wavelength, $\lambda = L$ (see also our previous work
with homogeneous substrates~\cite{PhysRevE.104.034801}).}
All our simulations are run on a bi-periodic square domain of size $L \times L$ with 
$L/h_0 = 512 $. \textcolor{red}{The numerical values in lattice Boltzmann units of the 
mean film thickness, dynamics viscosity and surface tension are, respectively, $h_0=1$, $\mu=1/6$
and $\gamma=0.01$.}
\textcolor{red}{To regularize the contact line divergence~\cite{huh1971hydrodynamic} we use a precursor layer thickness $h_{\ast}/h_0=0.07$ and a slip length $\delta/h_0 = 1$, 
a value which lies
within the weak/intermediate slip regime~\cite{peschka2019signatures,fetzer2007quantifying, munch2005lubrication} (we have tried also a value four times smaller, without observing 
any major qualitative change in the results).} 
The liquid film is initialized with a height field slightly perturbed around the mean value $h_0$, i.e. $h(\mathbf{x},0) = h_0 \left[1 + 0.1 \left(\sin\left(\frac{2\pi x}{L}\right)\sin\left(\frac{2\pi y}{L}\right)\right)\right]$.
Various wavelengths in the range \textcolor{red}{$\lambda/h_0 \in [64, 512]$}, and velocities,
$v_{\theta}/v_0 \in [0, 200]$, are considered for the wettability pattern, Eq.~(\ref{eq:sinetheta}). 
\textcolor{red}{A few comments on these ranges are noteworthy at this point. The wavelengths are obviously limited from above by the size of the computational domain. On the other extreme, too short wavelengths are prevented by
the energetic cost of accomodating droplets on too
small patches, since the local contact angle 
(that increases with decreasing $\lambda$~\cite{SuppMat}) would possibly exceed the largest substrate contact angle, such that the patterning 
itself would loose meaning. 
It has been, in fact, proven, both theoretically and experimentally, that 
for the confinement to be effective, the pattern characteristic size should 
not be much smaller than the spinodal length scale on the less wettable substrate
\cite{KarguptaJCP2002,KarguptaLangmuir2000,Nisato1999,Karim1998}.
We keep, therefore, $\lambda \stackrel{>}{\sim} \lambda_s \approx 70 h_0$.
The pattern speeds, on the other hand, are limited by the intrinsic LB bound of the lattice speed of sound
 \cite{succi}.}\\
%\textcolor{red}{Hereafter, we will omit 
%to write explicitly the division by $h_0$, $t_0$ and $v_0$ that will be, then, 
%surmised, unless else specified.}
\textcolor{red}{The Reynolds number, $Re$, is of the order of $Re \sim 10^{-2}$ on the 
static substrate and never exceeds the value $\approx 0.2$ in the time dependent case.}
In Fig.~\ref{fig:handtheta} we show $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e. the static case) and \textcolor{red}{$\lambda = 256 h_0$}\footnote{For a better visualization we take twice the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.\\

\noindent {\it Results.} We start our analysis by studying first how the rupture times depend on the parameters characterising the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (that is, when the free surface "touches" the substrate).

@@

We observe that $\tau_r$ grows linearly with $\lambda$, for short wavelengths, and quadratically for longer $\lambda$.
These facts can be explained, qualitatively, as follows. 
Let us first notice that the film rupture consists of two processes, each with its associated time scale: the growth of unstable interface perturbation with rate $t_{\theta}^{-1}$, and the retraction of liquid from more hydrophobic regions, occurring in a characteristic time $t_R$. For longer pattern wavelengths, the dewetting instability amplifies more slowly, such that it is reasonable to take it as the process determining the rupture time, $\tau_r \sim t_{\theta}$.

@@

@@

 \tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,

@@

    \includegraphics[width=0.45\textwidth]{Figure_2.pdf}

@@

%Interestingly, this phenomenology is qualitatively supported %also by the time-dependent case, $v_{\theta} = 20 v_0$ (reported %in Fig.~\ref{fig:model_rt} as orange stars), \textcolor{red}{but %for a smaller prefactor in the linear regime, due to the fact %that the pattern wave velocity is large enough to dominate over %$U_{\Theta}$ and sets the retraction speed, such that the %characteristic time becomes $\tau_R \sim \lambda/v_{\theta}$.}
%\begin{equation}\label{eq:taur_saturation}
% \tau_r \sim \tau_R \sim \frac{\lambda}{v_{\theta}} \propto %\frac{\lambda}{v_0} = \lambda\frac{t_0}{\lambda} \sim \mbox{const}.
We now focus on the long time dynamics and in particular on the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.

@@

    \includegraphics[width=0.45\textwidth]{Figure_3.pdf}

@@

      which is $2\left(\frac{L}{\lambda}\right)^2$. In the insets, we display snapshots of the stationary droplet states as grey-scale images

@@

On the stationary substrate, after rupture all the fluid accumulates in droplets centered at contact angle minima.
Consequently, as we see from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ in time\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}; the clusters are determined by means of an algorithm of Hoshen-Kopelman type~\cite{HK}.}, in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (reported as horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.

@@

\textcolor{red}{In the main panel of the same figure, we see that the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$

@@

droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to the decrease of the droplet volume,

@@

    \includegraphics[width=0.45\textwidth]{Figure_4.pdf}
    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$ (see text for details), for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
    The grey-scale insets supply snapshots of the film thickness field associated to the various curves.

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
    \includegraphics[width=0.45\textwidth]{Figure_5.pdf}
    \caption{MAIN PANEL: Rivulets life-times, $\tau_{\text{riv}}$, for various $\Gamma$ values.
    The dashed line is plotted as a guide to the eye to highlight the logarithmic dependence, in agreement with 
    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, computed along the rivulet axis, for three different $\Gamma$ values.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie file ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.

@@

If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than the transport due to the contact angle field and thus droplets from. However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state. \textcolor{red}{It appears, therefore, natural to consider the ratio of these two velocities, 

@@

Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, this should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration. Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore 

@@

The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$ \cite{PhysRevE.104.034801}, which yields

@@

\textcolor{red}{We would like to envisage a possible realization of a 

@@

This technology has effectively been used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}, it allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)

@@

Considering as a reference, for instance, the system studied in ~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 

>>>>

\bibstyle{apsrev}
\usepackage{bm}%
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{color}
\usepackage{xcolor}
\usepackage{physics}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}%
\expandafter\ifx\csname package@font\endcsname\relax\else
 \expandafter\expandafter
 \expandafter\usepackage
 \expandafter\expandafter
 \expandafter{\csname package@font\endcsname}%
\fi
\hyphenation{title}
\newcommand{\JH}[1]{\textcolor{blue}{ JH: #1}}
\newcommand{\REV}[1]{\textcolor{red}{#1}}
%\newcommand{\REV}[1]{#1}

@@

\begin{abstract}
\noindent Switchable and adaptive substrates emerged as valuable tools for controlling wetting and actuation of droplet motion. Here, we report a computational study of the dynamics of an unstable thin liquid film deposited on a switchable substrate, modeled with a space and time varying contact angle.
With a static pattern, all fluid is drained into droplets located around contact angle minima, whereas for a sufficiently large rate of wettability variation a state consisting of metastable rivulets is observed. 
A criterion discriminating whether rivulets occur or not is identified in terms of a single dimensionless parameter.
Finally, we show and explain theoretically how the film rupture times, droplet shape and rivulet life time depend on the pattern wavelength and speed.  
\end{abstract}

@@

\noindent {\it Introduction}. Wet surfaces and droplets are part of our every-day experience and of numerous industrial processes including coating, tribology, painting and printing, to name but a few~\cite{gross1980fluid,szeri2010fluid,doi:10.1146/annurev.fluid.31.1.347,DASILVASOBRINHO19991204,singh2010inkjet,Wijshoff2010}. 
Moreover, the continuously growing interest for lab-on-a-chip devices~\cite{C6LC00387G,Focke} as well as for printable electronics or printable photovoltaics~\cite{Luechinger_2008,RH22}, whose efficiency relies crucially on a precise control of material deposition upon (de-)wetting of liquid films, drew the attention to applications where the substrate is adaptive or switchable, i.e. it is not inert but responds dynamically to external stimuli or to the evolution of the coating liquid film itself~\cite{ButtEtAl_Langmuir2018}. 
Several realizations of switchable and adaptive substrates have been proposed~\cite{XiCSR2010}, involving smart materials such as polymer brushes~\cite{CohenStuartEtAl_NatMat2010}, thermal-responsive hydrogels~\cite{ChenEtAl_SM2010}, light-responsive molecules and microstructures~\cite{IchimuraEtAl_Science2000} or processes such as electrowetting~\cite{MugeleEtAl_JPCM2005}.
Probably the simplest, yet non-trivial, modelling of a switchable substrate can be realized by a space and time dependent wettability pattern~\cite{GrawitterStark1}. 
While a consistent body of theoretical/computational work was devoted to processes on static heterogeneous substrates, the time dependent case is still almost unexplored, with few exceptions focusing on single droplet spreading and sliding~\cite{GrawitterStark1,GrawitterStark2,ThieleHartmann} or limited to analysing the linear regime~\cite{suman2006dynamics}.

Here, we study by means of numerical simulations the full dewetting dynamics of a thin liquid film deposited on a substrate with a time varying wettability pattern, from rupture to the long time morphology. 
We identify two regimes where the rupture times grow with the pattern wavelength either linearly (on a static pattern) or attain a constant value (in the time dependent case), for short wavelengths, and approach a quadratic law as the wavelength increases. These observations are then explained theoretically.
We show that, by tuning the rate of change (the ``speed'') of the underlying pattern, one can control, to some extent, the dewetting morphology. In particular, for large enough pattern speeds, we detect a metastable state, where the film retracts into metastable rivulets, eventually breaking up into multiple droplets.
We introduce a control parameter to discriminate whether rivulets or just droplets (as in the static situation)
can be observed and propose a phenomenological argument to justify the logarithmic dependence of the rivulets life time on the pattern speed.\\

@@

\noindent {\it Method.} In order to simulate the dewetting dynamics on patterned, ``switchable'', substrates, we integrate numerically the thin-film equation~\cite{RevModPhys.69.931,RevModPhys.81.1131} 

@@

by means of a recently developed lattice Boltzmann (LB) scheme~\cite{PhysRevE.100.033313,PhysRevE.104.034801, Zitz2022}.

@@

The film pressure $p(\mathbf{x},t)$ consists of the sum of the Laplace and disjoining pressures, that is $p(\mathbf{x},t) = -\gamma \nabla^2 h - \Pi$ \REV{($\gamma$ is the surface tension)}~\footnote{\REV{The Laplace pressure is the product of the surface tension 
and the local curvature. Therefore, its full expression reads $p_L = -\gamma \nabla^2 h/\sqrt{1+|\nabla h|^2}$. 
In the lubrication regime, whereby $|\nabla h| \ll 1$, the denominator can be safely neglected 

@@

field}~\cite{Benet2014,juanes2018}}.

@@

\REV{\Pi(h,\theta) = \frac{2\gamma}{h_{\ast}}(1-\cos(\theta(\mathbf{x},t)))

@@

\REV{where $f(\xi)=\xi^{-3} - \xi^{-2}$}. $h_{\ast}$ is the height at which $\Pi$ vanishes and sets the precursor layer thickness~\cite{SuppMat}.
The time variation of the patterned substrate enters the model through the disjoining pressure, by making the contact angle space and time dependent, i.e.~$\theta = \theta(\mathbf{x},t)$.
\REV{We decorate the substrate with a checkerboard pattern, a common choice that generalizes the broken homogeneity of the stripes to two 
directions~\cite{Jalali2018,Nagayama2020,Das2020}.}

@@

% \REV{A possible laboratory experiment for the here mentioned approach consists of a light switchable substrate with a thin film cast on it~\cite{becker2003complex, IchimuraEtAl_Science2000}. 

@@

We fix the velocity direction to one diagonal, namely $\mathbf{v}_{\theta} = (v_{\theta x},v_{\theta y}) = v_{\theta}(1/\sqrt{2},-1/\sqrt{2})$ (we will return later on the importance of this choice), and we set $\theta_0 = 20^{\circ}$ and $\delta\theta=10^{\circ}$~\footnote{Since a typical velocity is such that $v_{\theta} \Delta t \ll \Delta x$ (in one time step $\Delta t$ the wave would travel a distance much smaller than a lattice spacing $\Delta x$),
the time update needs to be interpreted in an integer part sense, that is the pattern is shifted by $\Delta x$ every $1/v_{\theta x}$ $\Delta t$ (equivalently in $y$-direction).}.
\REV{Lengths and time scales will be expressed, respectively, in units of the mean film height, $h_0$ (which is constant in time, due to mass conservation), and of $t_0 = \frac{3\mu}{\gamma h_0^3 q_0^4}$, the inverse growth rate of the most unstable mode, whose wavenumber is $q_0$, of a spinodally dewetting film~\cite{Mecke_2005}. On a uniform substrate, with constant contact angle $\theta^{(u)}$, the wavenumber reads 
$(q^{(u)}_0)^2 = h_{\ast}^{-2}(1-\cos \theta^{(u)})f^{\prime}(h_0/h_{\ast})$~\cite{Mecke_2005,PhysRevE.100.023108}. In our patterned case, we define 

@@

    \includegraphics[width=0.4\textwidth]{Figure_1.png}

@@

    (Eq.~(\ref{eq:sinetheta}) with $v_{\theta}=0$), with lower (higher) values in light blue (yellow).

@@

Fig.~\ref{fig:handtheta} shows $h(\mathbf{x},t)$ (droplets) and $\theta(\mathbf{x})$ (color coded) for $v_{\theta} = 0$ (i.e. the static case) and \REV{$\lambda = 256 h_0$}~\footnote{For a better visualization double the domain length $L$ and periodically continue the image.}, in late stages of dewetting.
As expected, droplets form in regions of small contact angles (blue) while the regions of high contact angles (yellow) dewet.

\noindent {\it Results.} We first investigate how the rupture times depend on the parameters characterizing the wettability pattern, namely the wavelength of the contact angle variation, $\lambda$, and wave speed, $v_{\theta}$.
The film rupture time, $\tau_r$, is defined as the least $t$ such that $h(\mathbf{x},\tau_r)=h_{\ast}$ (i.e., when the free surface "touches" the substrate).

@@

We observe that $\tau_r$ grows linearly with $\lambda$ for short wavelengths and quadratically for longer $\lambda$.
These facts can be qualitatively explained as follows. 

@@

 %\tau_r \sim \tau_R \propto \frac{1}{U_{\Theta}}\lambda,

@@

 \tau_r \sim \tau_R \propto U_{\Theta}^{-1}\lambda,

@@

    \includegraphics[width=0.4\textwidth]{Figure_2.pdf}

@@

We now focus on the long time dynamics, the characterization of the dewetting morphologies and how they are affected by the speed of the wettability wave.

@@

    \includegraphics[width=0.4\textwidth]{Figure_3.pdf}

@@

      which is $2\left(\frac{L}{\lambda}\right)^2$. The napshots depict the stationary droplet states as grey-scale images

@@

On the stationary substrate, after rupture all fluid accumulates in droplets centered at contact angle minima.
Consequently, as seen from the inset of Fig.~\ref{fig:clusters_v0_sine}, where we plot the number of droplets $N(t)$ vs. time~\footnote{A droplet is identified by the set (``cluster'') of points, in the plane, constituting each of the connected components of the set $\{\mathbf{x} \in [0,L]^2 | h(\mathbf{x},t) \geq h_{\ast}$\}.
% the clusters are determined by means of a Hoshen-Kopelman algorithm~\cite{HK}.
},
in the steady state ($t \gg t_0$) $N(t)$ attains the value $N_{\infty} = 2(L/\lambda)^2$ (horizontal lines), which equals the minima of Eq.~(\ref{eq:sinetheta}), for $v_{\theta}=0$.

@@

\REV{In the main panel of the figure, the height fluctuations $\Delta h(t) = \max_{\mathbf{x}}\{h(\mathbf{x},t)\}-\min_{\mathbf{x}}\{h(\mathbf{x},t)\}$

@@

droplets are essentially monodisperse), decreasing with the pattern wavelength (as expected, due to a decreasing droplet volume,

@@

    \includegraphics[width=0.4\textwidth]{Figure_4.pdf}
    \caption{Time evolution of the second order Minkowski structure metric, $q_2(t)$, for different $\Gamma$ values, on a substrate with pattern wavelength $\lambda=256 h_0$.
    The grey-scale insets supply snapshots of the corresponding film thickness fields.}

@@

In order to better characterize the various morphologies we apply the theory of Minkowski's functionals. 
In particular, we employ the second order Minkowski structure metric, $q_2$~\cite{doi:10.1063/1.4774084, Schaller2020}, which can be computed from a Voronoi tessellation of the set of discrete points $(x_i, y_i)$ on the 2D lattice, such that the height field lies above a certain threshold\footnote{The expression is: $q_2 = \frac{1}{N}\sum_j \frac{1}{P_j}|\sum_k L^{(j)}_k e^{2i\phi^{(j)}_k}|$, where the inner sum runs over the edges of length $L^{(j)}_k$, of the $j$-th Voronoi cell, whose perimeter is $P_j$, and $\phi^{(j)}_k$ is the polar angle of the normal to the $k$-th edge. 
The outer sum represents an ensemble average over the $N$ points in the set.}. The $q_2$ metric quantifies the degree of anisotropy of the dewetting morphology, so it takes relatively large values if the structures formed display a preferential direction. 
Measuring $q_2$ then enables us to clearly distinguish between the formation of droplets and rivulets: much larger $q_2$ values are attained for the latter type of structure, as we can see in Fig.~\ref{fig:msm_q2}.
%Our simulations with $\lambda = L/2$ show a clear difference between $v_{\theta} \leq 0.1v_0$ and $v_{\theta}>0.1v_0$  in $q_2$.
We observe, on the other hand, that such rivulets are metastable and eventually break up into droplets, as indicated by the collapse of $q_2$ at later times. 
Notice, though, that the $q_2$ signal for any $v_{\theta} >0$ always stays above the one for the static case, suggesting that even the smallest pattern velocity introduces a sizeable deformation of the spherical cap shape.
\begin{figure}
    \centering
    % \includegraphics[width=0.45\textwidth]{Figures/Lig_stab_inset_delta_h.pdf}
    \includegraphics[width=0.4\textwidth]{Figure_5.pdf}
    \caption{MAIN PANEL: Rivulet life-times, $\tau_{\text{riv}}$, for various $\Gamma$.
    The dashed line is a guide to the eye to highlight the logarithmic dependence, in agreement with 
    the theoretical prediction, Eq.~(\ref{eq:rivlt}).
    INSET: Height fluctuations, $\Delta h(t)$, vs time, along the rivulet axis, for three different $\Gamma$.
    }
    \label{fig:stab_ligs_lam2}
\end{figure}
The breakup is the result of a varicose mode of the rivulet~\cite{doi:10.1063/1.3211248, PhysRevE.77.061605}, whose wavelength is $\approx\lambda$, such that only $N_{\infty}/2$ droplets are counted after breakup. 
These droplets show a peculiar dynamics, characterized by a periodic sequence of spreading and retraction, driven by the pattern, that we dub "pumping state"~\footnote{see movie ligament\_formation\_and\_breakup.mp4 in~\cite{SuppMat}.}.

@@

\REV{If $U_{\Theta}$ is large as compared to $v_{\theta}$, the film retraction is faster than 
the local contact angle variation and thus droplets form.}
%the transport due to the contact angle field and thus droplets from. 
However, if $v_{\theta}$ is larger than $U_{\Theta}$, then the retracting film has too little time to form droplets and ends up in the metastable rivulet state. \REV{It appears, therefore, natural to consider the ratio of these two velocities, 

@@

Heuristically speaking, this means that, if we evaluate $\Delta h(t)$ restricted on the rivulet axis, it should grow exponentially (with a certain growth rate $\alpha \propto t_0$) only when the system is in the unstable configuration. Namely, $\Delta h(t)/\Delta h_0 \propto e^{\alpha t}$ (see inset of Fig.~\ref{fig:stab_ligs_lam2}) with a prefactor proportional to the time spent by the rivulet in such a configuration, which goes as $\sim \lambda/v_{\theta}$, therefore 

@@

The rivulet life-time can be seen as the rupture time of the structure along its axis, hence such that $\Delta h (\tau_{\text{riv}}) \sim h_0$~\cite{PhysRevE.104.034801}, which yields

@@

\REV{We envisage a possible realization of a 

@@

This technology was effectively used for thin film experiments and additive manufacturing~\cite{doi:10.1021/jp301092y, doi:10.1126/science.aax8760}. It allows for fast temporal modulations of the optical signal (with frequencies up to $\approx 16 \, \text{kHz}$)

@@

Considering as a reference, for instance, the system studied in~\cite{becker2003complex,PhysRevLett.99.114503}, namely a $\sim 4 \, \text{nm}$ thick film 
